### 2019-07-11 submission/manuscript.Rmd
Additional edits before sharing with Pat

**PPI-treatment did not promote susceptibility to *C. difficile infection* in mice**
Next, we examined whether high dose PPI treatment altered susceptibility to *C. difficile* infection in mice. After 7 days of PPI treatment or 1 day after clindamycin treatment, mice were challenged with *C. difficile* 630 spores. While all 4 of the clindamycin-treated mice were colonized with *C. difficile*, all of the PPI-treated mice were resistant to *C. difficile* colonization (Figure 2A). Interestingly, only 1 cage of the PPI and clindamycin-treated mice were colonized, while the other cage was not (Figure 2A). PCoAs of the bacterial communities from the 3 groups of mice after *C. difficile* challenge, showed that the greatest shifts in bacterial communities occurred in the clindamycin-treated mice (Figure 2B). Within 5 days, all of the mice had cleared *C. difficile*, suggesting there was no difference in rate of clearance for the mice that were initially colonized with *C. difficile* despite continuing to receive PPI treatment throughout the course of the experiment. Our results suggest that PPI treatment alone had no effects on bacterial community resistance to *C. difficile* colonization in mice. Instead most of the differences between our 3 treatment groups appear to be driven by clindamycin administration and included decreased *Alistipes*, *Barnesiella*, Porhyromonadaceae, Ruminococcaceae, taxa previously found to be altered in clindamycin-treated mice that were challenged with *C. difficile* [@Schubert2015]. These findings demonstrated that high dose PPI treatment did not promote susceptibility to *C. difficile* colonization.

@@

Our findings that PPI treatment had minimal impact on the fecal microbiome were comparable to another PPI mouse study that indicated PPIs have more of an effect on the small intestinal microbiota compared to the fecal microbiota [@Yasutomi2018]. The same group demonstrated PPI treatment increased the stomach pH in mice [@Yasutomi2018], which could improve survival of bacteria passing through the stomach. We did not find significant changes for the taxa observed to be significantly impacted by PPI use in human studies. However, 3 of the families that typically increase were absent or at low abundance in the mice from the beginning and may be one contributing factor to our observation that PPIs had no significant effects on bacterial taxa. Additionally, some of the significance of PPI associations in human interventional trials appears to be driven by a handful of specific taxa with overall differences on a PCoA plot difficult to distinguish [@Freedberg2015, @Imhann2015]. Also, most of the studies to date compared different individuals (PPI users versus non-users) or the same individuals before and after treatment [@Naito2018; @Imhann2017], unfortunately these limited microbiota snapshots may be skewed by day-to-day microbiota variations [@LlornsRico2019] that would be revealed with more frequent longitudinal sampling.   
ALthough there have been a few *C. difficile* mouse model studies that have demonstrated PPIs have some effect on CDIs with or without additional antibiotic treatment[@Kaur2007; @kaur2011effect; @hung2015proton], there were some key methodological differences between the previous studies and our own. One group administered 0.5 mg/kg of the PPI lansoprazole daily for 2 weeks to mice and then challenged with *C. difficile* demonstrated that PPI treatment alone resulted in detectable *C. difficile* in the stool 1 week after challenge, but also showed there was detectable *C. difficile* in mice not treated with antibiotics [@Kaur2007; @kaur2011effect]. The presence of *C. difficile* in mice that were not treated with either antibiotics or PPIs could be attributed to the higher dose of vegetative cells (10^8^ CFU) used to challenge the mice and may partly explain the observation of *C. difficile* in PPI-treated mice [@Kaur2007; @kaur2011effect]. We have previously shown mice from our colony that were not given antibiotics were resistant to *C. difficile* 630 when challenged with 10^3^ spores [@Jenior2017]. The other group also used a higher dose of 10^6^ spores or 10^7^ vegetative cells to challenge antibiotic-treated mice or mice treated with both the PPI esomeprazole (40mg/kg dose) for 2 days and antibiotics and demonstrated the antibiotic/PPI-treated mice developed more severe CDIs [@hung2015proton].  
Our study extended previous work examining PPIs and *C. difficile* in mice to examine the contribution of the intestinal microbiota. We found the PPI omeprazole had minimal impact on the murine intestinal microbiota and in contrast to previous work, PPIs did not promote *C. difficile* colonization. 16S rRNA sequencing suggested that *Streptococcus* and *Enterococcus* are rare genera in our C57BL/6 mouse colony. These genera could be important contributors to the associations between PPIs and CDIs in humans, and could be a contributing factor to why we observed that PPI treatment had no effect on *C. difficile* colonization in our CDI mouse model. There could also be differences at the species level for taxa that are altered with PPIs in humans compared to mice. Other murine starting communities or the addition of human PPI treatment-associated strains may be needed to determine whether PPIs can alter taxa and subsequently promote CDIs.  

>>>>

**PPI-treatment did not promote susceptibility to *C. difficile infection* in mice**
Next, we examined whether high dose PPI treatment altered susceptibility to *C. difficile* infection in mice. After 7 days of PPI treatment or 1 day after clindamycin treatment, mice were challenged with *C. difficile* 630 spores. While all 4 of the clindamycin-treated mice were colonized with *C. difficile*, all of the PPI-treated mice were resistant to *C. difficile* colonization (Figure 2A). Interestingly, only 1 cage of the PPI and clindamycin-treated mice were colonized, while the other cage was resistant (Figure 2A). PCoAs of the bacterial communities from the 3 groups of mice after *C. difficile* challenge, showed that the greatest shifts in bacterial communities occurred in the clindamycin-treated mice (Figure 2B). Within 5 days, all of the mice had cleared *C. difficile*, suggesting there was no difference in rate of clearance for the mice that were initially colonized with *C. difficile* despite continuing to receive PPI treatment throughout the course of the experiment. Our results suggest that PPI treatment alone had no effects on bacterial community resistance to *C. difficile* colonization in mice. Instead most of the differences between our 3 treatment groups appear to be driven by clindamycin administration and included decreased *Alistipes*, *Barnesiella*, Porhyromonadaceae, Ruminococcaceae, taxa previously found to be altered in clindamycin-treated mice that were challenged with *C. difficile* [@Schubert2015]. These findings demonstrated that high dose PPI treatment did not promote susceptibility to *C. difficile* colonization.

@@

Our findings that PPI treatment had minimal impact on the fecal microbiome were comparable to another PPI mouse study that indicated PPIs had more of an effect on the small intestinal microbiota compared to the fecal microbiota [@Yasutomi2018]. The same group demonstrated PPI treatment increased the stomach pH in mice [@Yasutomi2018], which may improve survival of bacteria passing through the stomach. We did not find significant changes for the taxa observed to be significantly impacted by PPI use in human studies. However, 3 of the families that typically increase were absent or at low abundance in the mice from the beginning and may be one contributing factor to our observation that PPIs had no significant effects on bacterial taxa. Additionally, some of the significance of PPI associations in human interventional trials appears to be driven by a handful of specific taxa with overall differences on a PCoA plot difficult to distinguish [@Freedberg2015, @Imhann2015]. Also, most of the studies to date compared different individuals (PPI users versus non-users) or the same individuals before and after treatment [@Naito2018; @Imhann2017], unfortunately these limited microbiota snapshots may be skewed by day-to-day microbiota variations [@LlornsRico2019] that would be revealed with more frequent longitudinal sampling.   
ALthough there have been a few *C. difficile* mouse model studies that have demonstrated PPIs have some effect on CDIs with or without additional antibiotic treatment[@Kaur2007; @kaur2011effect; @hung2015proton], there were some key methodological differences between these studies and our own. One group administered 0.5 mg/kg of the PPI lansoprazole daily for 2 weeks to mice and then challenged with *C. difficile* demonstrated that PPI treatment alone resulted in detectable *C. difficile* in the stool 1 week after challenge, but also showed there was detectable *C. difficile* in mice not treated with antibiotics [@Kaur2007; @kaur2011effect]. The presence of *C. difficile* in mice that were not treated with either antibiotics or PPIs could be attributed to the higher dose of vegetative cells (10^8^ CFU) used to challenge the mice and may partly explain the observation of *C. difficile* in PPI-treated mice [@Kaur2007; @kaur2011effect]. We have previously shown mice from our colony that were not given antibiotics were resistant to *C. difficile* 630 when challenged with 10^3^ spores [@Jenior2017]. The other group also used a higher dose of 10^6^ spores or 10^7^ vegetative cells to challenge antibiotic-treated mice or mice treated with both the PPI esomeprazole (40mg/kg dose) for 2 days and antibiotics and demonstrated the antibiotic/PPI-treated mice developed more severe CDIs [@hung2015proton].  
Our study extended previous work examining PPIs and *C. difficile* in mice to examine the contribution of the intestinal microbiota. We found the PPI omeprazole had minimal impact on the murine intestinal microbiota and in contrast to previous work, PPIs did not promote *C. difficile* colonization. 16S rRNA sequencing suggested that *Streptococcus* and *Enterococcus* are rare genera in our C57BL/6 mouse colony. These genera could be important contributors to the associations between PPIs and CDIs in humans, and could be a contributing factor to our observation that PPI treatment had no effect on *C. difficile* colonization in our CDI mouse model. There could also be differences at the species level for taxa that are altered with PPIs in humans compared to mice. Other murine starting communities or the addition of human PPI treatment-associated strains may be needed to determine whether PPIs can alter taxa and subsequently promote CDIs.  
